# This file describes the analysis of the termite phylogenetic community structure in the Amazonian Forest



```{r Load Packages and functions}

library(ape)
library(picante)
library(tree)
library(reshape)
library(raster)
library(maptools)

source("phyloBeta.R")

```


```{r import data}

folder<-"Shapefiles/"

brazil<-readShapeSpatial(paste(folder,"BR_Contorno.shp",sep=""))
biomes<-readShapeSpatial(paste(folder,"bioma.shp",sep=""))

isoptera<-read.csv("Isoptera.new.csv",row.names=1)

#Removes repeated records of the same species in the same subplot (5x2 quadrant)

isoptera<-isoptera[!duplicated(isoptera[,c("Taxon","Location","Grid","Trail","Plot","Distance_beginning","Side")]),]

env<-read.csv("environment.csv")

geo.dist<-dist(cbind(env$LAT,env$LONG,env$Alt*0))

env.dist<-dist(cbind(env$Clay))


```


### Uses taxonomic information to build the basic phylo tree

```{r construct taxonomic tree}

sp.ls<-unique(isoptera[,c("Taxon","Genus","Subfamily","Family","Order")])

sp.ls.int<-as.matrix(data.frame(lapply(sp.ls,as.integer)))

sp.ls.int[is.na(sp.ls.int)]<-0

# Calculates the taxonomic distance among all species pairs (the code can be improved - get rid of the for loop )

sp.ls.dist<-(sp.ls.int[,1])%*%t(sp.ls.int[,1]*0+1)!=(sp.ls.int[,1]*0+1)%*%t(sp.ls.int[,1])

for(i in 2:ncol(sp.ls.int)){
	sp.ls.dist<-
		as.matrix(sp.ls.dist)+
		as.matrix((sp.ls.int[,i])%*%t(sp.ls.int[,i]*0+1)!=(sp.ls.int[,i]*0+1)%*%t(sp.ls.int[,i]))
}

dimnames(sp.ls.dist)<-list(sp.ls[,1],sp.ls[,1])

write.table(sp.ls.dist,"taxodist.csv")

sp.ls.dist<-as.dist(sp.ls.dist)

termite.taxophylo<-as.phylo(hclust(sp.ls.dist))

write.tree(termite.taxophylo,"termite.taxophylo.tre")

# cophenetic(termite.taxophylo) returns the original taxonomic distance

```


```{r }

plot(termite.taxophylo,cex = .5)

```


```{r create response variables}

attach(isoptera)

#write.csv(unique(data.frame(Location,Grid,Trail,Plot))[-c(1:148),],"jau.csv")

loc<-factor(paste(Location,Grid,Trail,Plot),levels=paste(env$Location,env$Grid,env$Trail,env$Plot))

loc.split<-unique(data.frame(env$Location,env$Grid,env$Trail,env$Plot))

#sum(loc!=paste(Location,Grid,Trail,Plot)) #Must be zero

termite.plot<-tapply(Frequency,list(loc,Taxon),sum)
termite.plot[is.na(termite.plot)]<-0

termite.plot.PA<-(termite.plot>0)+0 # Easier and faster than ifelse(termite.plot>0,1,0)

### Simiarity indexes

sor.dist<-vegdist(termite.plot.PA,"bray",binary = TRUE)

detach(isoptera)

```


```{r}

plot(biomes[1,])
points(env$LONG,env$LAT,pch=21,bg=adjustcolor(1,alpha=.1),col=0,lwd=.1,cex=2)

```


``` {r }

dist.PDB<-matrix(NA,nrow(termite.plot),nrow(termite.plot))

for(i in 1:nrow(dist.PDB)){
	for(j in 1:nrow(dist.PDB)){
		
if(j<i&sum(termite.plot[i,])>1&sum(termite.plot[j,])>1){		

termite.phylocom<-melt(termite.plot[c(i,j),])[,c(1,3,2)]
termite.phylocom<-termite.phylocom[termite.phylocom[,2]>0,]
PDcalc(termite.taxophylo,termite.phylocom)
dist.PDB[i,j]<-phyloBeta(termite.taxophylo,termite.phylocom)$PD.beta
		
	}
	
		}
		
	}

dist2.PDB<-as.dist(dist.PDB)

dim(dist2.PDB)

length(dist2.PDB)
length(dist)

plot(dist2.PDB~dist,pch=21,col=0,bg=adjustcolor(2,alpha=.1))

```



#Regression Tree

```{r}

PCA.term<-prcomp(termite.plot)$x[,1]

reg.tree<-tree(PCA.term~LAT+Clay+Altitude+LONG+P,data=env)

regtree$frame$splits<-ifelse(regtree$frame$splits!=0,"","")

plot(regtree,lwd=2)
text(regtree,cex=.8,adj=1.2)


plot(PCA.term~env$LAT,col=env$Location)

```


```{r}


ENSpie<-(1/(1-apply(termite.plot,1,PIE)))

ENSpie[ENSpie<0|ENSpie>1000]<-NA

plot(tree(ENSpie~LONG+Altitude,data=env))
text(tree(ENSpie~LONG+Altitude,data=env))


data.frame(ENSpie,env[,1:3])

[is.na(env$Clay),]

summary(lm(ENSpie~LAT+LONG+Clay+Altitude,data=env))
env




```



```{r Phylogenetic Analysis}

## PhyloBetaDiver

ufrac.termite<-unifrac(termite.plot.PA,termite.taxophylo)
psor.termite<-phylosor(termite.plot.PA,termite.taxophylo)
betaMNTD.termite<-comdistnt(termite.plot.PA,cophenetic(termite.taxophylo))

```

```{r Analysis-NMDSaxes}
NMDS.sor<-metaMDS(sor.dist,k=2)
NMDS.ufrac<-metaMDS(ufrac.termite,k=2)
NMDS.psor<-metaMDS(psor.termite,k=2)
NMDS.betaMNTD<-metaMDS(betaMNTD.termite,k=2)

summary(lm(scores(NMDS.sor)[,1]~env$LAT+env$LONG+env$Clay))
summary(lm(scores(NMDS.ufrac)[,1]~env$LAT+env$LONG+env$Clay))
summary(lm(scores(NMDS.psor)[,1]~env$LAT+env$LONG+env$Clay))
summary(lm(scores(NMDS.betaMNTD)[,1]~env$LAT+env$LONG+env$Clay))

summary(lm(scores(NMDS.sor)[,1]~env$LAT+env$LONG+env$Clay))
summary(lm(scores(NMDS.ufrac)[,1]~env$LAT+env$LONG+env$Clay))
summary(lm(scores(NMDS.psor)[,1]~env$LAT+env$LONG+env$Clay))
summary(lm(scores(NMDS.betaMNTD)[,1]~env$LAT+env$LONG+env$Clay))

```


```{r analyses using raw data (RDA)}


```


Following the suggestion of Tuomisto and Ruokolainen (2006), we conducted multiple regressions using the distance matrices (Sor,PSor,Ufrac and betaNMDS against DIST and envDIST). 

 We fit GLM models with binomial distribution and a log link for testing the association of taxonomic and phylogenetic similarity against geographical distance, except for betaMNTD that is not a probabilistic measure. For betaMNTD, the model was fit with a poisson distribution and a log link.

``` {r analyses using distance based data (mantel and regression in distances)}

#regressions using similarity matrices
sor.glm<-glm(sor.dist~geo.dist+env.dist,family=binomial(link="log"))
ufrac.glm<-glm(ufrac.termite~geo.dist+env.dist,family=binomial(link="log"))
psor.glm<-glm(psor.termite~geo.dist+env.dist,family=binomial(link="log"))
betaMNTD.glm<-glm(betaMNTD.termite~geo.dist+env.dist,family="poisson")

summary(sor.glm)
summary(ufrac.glm)
summary(psor.glm)
summary(betaMNTD.glm)











#Mantel tests
mantel(log(sor.dist+0.001),geo.dist,na.rm=TRUE)
mantel(log(ufrac.termite+0.001),geo.dist,na.rm=TRUE)
mantel(log(1-psor.termite+0.001),geo.dist,na.rm=TRUE)
mantel(log(betaMNTD.termite),geo.dist,na.rm = TRUE)

mantel(log(sor.dist+0.001),env.dist,na.rm=TRUE)
mantel(log(ufrac.termite+0.001),env.dist,na.rm=TRUE)
mantel(log(1-psor.termite+0.001),env.dist,na.rm=TRUE)
mantel(log(betaMNTD.termite+0.001),env.dist,na.rm = TRUE)

#Partial Mantel tests

mantel.partial(log(sor.dist+0.001),geo.dist,env.dist,na.rm=TRUE)
mantel.partial(log(ufrac.termite+0.001),geo.dist,env.dist,na.rm=TRUE)
mantel.partial(log(1-psor.termite+0.001),geo.dist,env.dist,na.rm=TRUE)
mantel.partial(log(betaMNTD.termite),geo.dist,env.dist,na.rm = TRUE)

mantel.partial(log(sor.dist+0.001),env.dist,geo.dist,na.rm=TRUE)
mantel.partial(log(ufrac.termite+0.001),env.dist,geo.dist,na.rm=TRUE)
mantel.partial(log(1-psor.termite+0.001),env.dist,geo.dist,na.rm=TRUE)
mantel.partial(log(betaMNTD.termite+0.001),env.dist,geo.dist,na.rm = TRUE)



```


```{r plots}


newdata=list(geo.dist=seq(0,15,by=.1))

plot(1-sor.dist~geo.dist,pch=21,bg="dark grey",col=0,xlab="Geographical Distance",ylab="Sorensen index",ylim=c(0,1))
lines(newdata,1-predict(sor.glm,newdata),col=2,lwd=2)

plot(1-ufrac.termite~geo.dist,pch=21,bg="dark grey",col=0,xlab="Geographical Distance",ylab="Unifrac index",ylim=c(0,1))
lines(newdata,1-predict(sor.glm,newdata),col=2,lwd=2)

plot(psor.termite~geo.dist,pch=21,bg="dark grey",col=0,xlab="Geographical Distance",ylab="PhyloSor index",ylim=c(0,1))
lines(newdata,1-predict(sor.glm,newdata),col=2,lwd=2)

plot(betaMNTD.termite~geo.dist,pch=21,bg="dark grey",col=0,xlab="Geographical Distance",ylab="betaMNTD")
lines(newdata,1-predict(sor.glm,newdata),col=2,lwd=2)


plot(1-ufrac.termite~env.dist)
plot(psor.termite~env.dist)
plot(betaMNTD.termite~env.dist)



plot(scores(NMDS.betaMNTD),col=as.factor(levels(loc)))

plot(scores(NMDS.sor),pch=21,col="dark grey",bg=heat.colors(1000)[round(decostand(env$Clay,"range",na.rm=T),3)*999+1])
plot(scores(NMDS.betaMNTD),pch=21,col="dark grey",bg=heat.colors(1000)[round(decostand(env$Clay,"range",na.rm=T),3)*999+1])



mpd(termite.plot.PA,cophenetic(termite.taxophylo))
#mpd(termite.plot.PA,as.matrix(sp.ls.dist)) #The same

ses.mpd<-ses.mpd(termite.plot.PA,cophenetic(termite.taxophylo))

```


```{r Trait analysis}

# Using functional traits (trophic group)



# Using ecological traits (like in Helmus et al. 2007a,b)



```

