---
title: "Phylogenetic and functional diveristy of termites"
author: "CSDambros"
date: "05/19/2014"
dates: "--" 
output:
  pdf_document:
    number_sections: yes
    toc: yes
    fig_caption: true
abstract: |
  This document describes in detail all the analyses performed for the phylogenetic and functional diversity of termites in response to ant predation, density of trees, P, and N content. It provides an R script follwing the analyses step by step, from downloading the data from public and permanent repositories to creating graphs and statistical tables. These are some of the ecological questions we tried to answer in this study:
  
  * Termites have a strong turnover along P content, but their densities are strongly associated with the density of predatory ants. 
  
  * Why are termite species replacing each other along the environmental gradients?
  
  * Does the phylogenetic community structure indicates environmental filtering or competition to sort the species?
  
  * How ant density affects the co-occurrence of termite species? Are some termites more vulnerable to ant predation?
---

\newpage


```{r setup,echo=FALSE,results='hide'}
library(knitr)

opts_chunk$set(warning=FALSE,message=FALSE)

```


# Load required packages

Some of the analyses run in our study were coded along the text, some analyses were in pre built packages, and some were transformed in functions, which are publicly available on-line.

The following lines load the required libraries, and download and source the necessary functions


```{r Load_Packages_and_functions, results='hide'}

library(ape)
library(picante)
library(tree)
library(vegan)
library(FD)
source(url("http://figshare.com/download/file/1672057/1")) # poncho function for figures
```

```{r,echo=FALSE}
#source(url("http://onlinelibrary.wiley.com/store/10.1111/j.1466-8238.2009.00490.x/asset/supinfo/GEB_490_sm_Appendix_S2.R?v=1&s=5363744bf3127ab77f2438bc5151bca2b30e3d96"))
```

# Import data

Similarly to the functions, we made our datasets publicly available on-line. They are downloaded using the next chunk of code. If you have problems running these lines, you might have an internet connection problem. Try to download the files to your directory using the provided links, and then read the files directly from your folder.


```{r import_data}

isoptera<-read.csv("Isoptera.new.csv",row.names=1) # termite data
traits<-read.csv("species.traits2.csv",row.names=1) # termite trait data
env<-read.csv("environment.csv") # envionmental variables, including ant data
termite.phylo<-read.nexus("termitephylo.fix.nex")# termite phylogeny

```
To facilitate handling objects and to reduce the number of objects in the workspace, we incorporated the trait information by species into the long format table with the termite record by colony 

```{r merge_and_clean}

# Include trait data into the termite data
isoptera$FG2<-traits$FG1[match(isoptera$Taxon,rownames(traits))]
isoptera$DEF<-traits$Maj[match(isoptera$Taxon,rownames(traits))]
```

To avoid counting the same termite colony twice, rows with exactly the same information were deleted. Therefore, colonies of the same species found in the same 5x2 plot were counted just once.

```{r cutting_duplicates}
isoptera<-isoptera[!duplicated(isoptera[,c("Taxon","Location","Grid","Trail","Plot","Distance_beginning","Side")]),]
iso.RFAD<-isoptera[isoptera$Grid=="RFAD",] #To be fixed and deleted
```
```{r selectproper, echo=FALSE,results='hide'}
# To be fixed and deleted
iso.RFAD<-isoptera[isoptera$Grid=="RFAD",]
env.RFAD<-cbind(env[env$Grid=="RFAD",1:7],read.csv("ducke.env.csv"),read.csv("env.RFAD.csv"))
```

```{r, echo=FALSE}
# scholtz<-read.csv("scholtz.csv")
# scholtz.g<-do.call(rbind,strsplit(as.character(scholtz$spp)," "))[,1]
# 
# scholtz<-read.csv("genuses.csv")
# 
# traits<-cbind(traits,scholtz[match(isoptera$Genus[match(rownames(traits),isoptera$Taxon)],scholtz$genus),])
# 
# write.csv(traits,"species.traits2.csv")
```

```{r test, echo=FALSE,eval=FALSE}

###

# summary(lm(env.RFAD$P~env.RFAD$lit.P_g.Kg))
# summary(lm(env.RFAD$P~env.RFAD$Lit.depth))
# summary(lm(env.RFAD$P~env.RFAD$LIS.1))
# summary(lm(env.RFAD$P~env.RFAD$LIS.2))
# summary(lm(env.RFAD$P~env.RFAD$LIS.3))
# summary(lm(env.RFAD$P~env.RFAD$BELT.1))
# summary(lm(env.RFAD$P~env.RFAD$BELT.2))
# summary(lm(env.RFAD$P~env.RFAD$BELT.3))
# summary(lm(env.RFAD$P~env.RFAD$cla_1))
# summary(lm(env.RFAD$P~env.RFAD$cla_2))
# summary(lm(env.RFAD$P~env.RFAD$cla_3))
# summary(lm(env.RFAD$P~env.RFAD$cla_4))
# summary(lm(env.RFAD$P~env.RFAD$cla_5))
# summary(lm(env.RFAD$P~env.RFAD$cla_6))
# summary(lm(env.RFAD$P~env.RFAD$cla_7))
# summary(lm(env.RFAD$P~env.RFAD$cla_8))
# summary(lm(env.RFAD$P~nmds.tree))
# 
# 
# summary(lm(env.RFAD$termite.N~I(env.RFAD$cla_2/env.RFAD$cla_1)))
# 
# 
# summary(lm(env.RFAD$termite.NMDS1~familyP.plot))
# 
# summary(lm(env.RFAD$lit.P_g.Kg~familyP.plot))
# 
# summary(lm(env.RFAD$termite.NMDS1~env.RFAD$lit.Mg_g.Kg))
# 
# summary(lm(env.RFAD$termite.NMDS1~env.RFAD$Lit.depth))
# summary(lm(env.RFAD$termite.NMDS1~env.RFAD$LIS.1))
# summary(lm(env.RFAD$termite.NMDS1~env.RFAD$LIS.2))
# 
# summary(lm(env.RFAD$termite.NMDS1~env.RFAD$LIS.3))
# 
# summary(lm(env.RFAD$termite.NMDS1~env.RFAD$BELT.1))
# summary(lm(env.RFAD$termite.NMDS1~env.RFAD$BELT.2))
# 
# summary(lm(env.RFAD$termite.NMDS1~env.RFAD$BELT.3))
# 
# summary(lm(env.RFAD$termite.NMDS1~env.RFAD$cla_1))
# summary(lm(env.RFAD$termite.NMDS1~env.RFAD$cla_2))
# summary(lm(env.RFAD$termite.NMDS1~env.RFAD$cla_3))
# summary(lm(env.RFAD$termite.NMDS1~env.RFAD$cla_4))
# summary(lm(env.RFAD$termite.NMDS1~env.RFAD$cla_5))
# summary(lm(env.RFAD$termite.NMDS1~env.RFAD$cla_6))
# summary(lm(env.RFAD$termite.NMDS1~env.RFAD$cla_7))
# summary(lm(env.RFAD$termite.NMDS1~env.RFAD$cla_8))
# 
# summary(lm(env.RFAD$termite.NMDS1~I(env.RFAD$BELT.3/env.RFAD$BELT.1)))
# 
# poncho(termite.plot.PA[!is.na(env.RFAD$LIS.3),],env=env.RFAD$LIS.3[!is.na(env.RFAD$LIS.3)],col=as.integer(TG.RFAD)+1,file="phyloxPA",ylab.bottom = expression("P "(mg/dm^3)),xlab="Sampling units",col.gradient = "tomato",border="black")
# 
# poncho(termite.plot.PA[!is.na(env.RFAD$BELT.3),],env=env.RFAD$BELT.3[!is.na(env.RFAD$BELT.3)],col=as.integer(TG.RFAD)+1,file="phyloxPA",ylab.bottom = expression("P "(mg/dm^3)),xlab="Sampling units",col.gradient = "tomato",border="black")
# 
# 
# poncho(termite.plot.PA[!is.na(env.RFAD$BELT.3),],env=env.RFAD$BELT.3[!is.na(env.RFAD$BELT.3)],col=as.integer(TG.RFAD)+1,file="phyloxPA",ylab.bottom = expression("P "(mg/dm^3)),xlab="Sampling units",col.gradient = "tomato",border="black")

```


```{r, echo=FALSE}

#Correct old phylo tree (Ruptitermes was in the wrong place)

# termite.phylo<-read.nexus("termitephylo.nex")
# 
# plot(termite.phylo)
# nodelabels()
# 
# rupti<-extract.clade(termite.phylo,89)
# phylo.wrupti<-drop.tip(termite.phylo,c(7,8))
# 
# termite.phylo<-bind.tree(phylo.wrupti,rupti,114)
# 
# plot(termite.phylo)
# nodelabels()
# 
# termite.phylo$edge.length[c(1,7,12)]<-termite.phylo$edge.length[c(1,7,12)]-1
# termite.phylo$edge.length[75]<-termite.phylo$edge.length[75]+1
# 
# write.nexus(termite.phylo,file="termitephylo.fix.nex")

```

# Extract data from tables

The termie data were stored in the long format. This means that each colony found is represented by a row in the spreadsheet. The species the colonies belong to are attribute of the colonies, so there is a column designating this attribute. All the analyses used in our study require a short table format, where species are in columns and the location they were found are rows. The following lines extract the information from the long format table, and rearange this information in the proper format for analyses. Finally, vectors with the species traits are created. These vectors are in the same order as the species in the new table. 

```{r extract_data_from_tables}

attach(iso.RFAD)

#Create factor with complete location
loc<-factor(paste(Location,Grid,Trail,Plot),
	levels=paste(env.RFAD$Location,env.RFAD$Grid,env.RFAD$Trail,env.RFAD$Plot))

#Just record the levels in the split format
loc.split<-unique(data.frame(env.RFAD$Location,
	env.RFAD$Grid,env.RFAD$Trail,env.RFAD$Plot))

#sum(loc!=paste(Location,Grid,Trail,Plot)) #Must be zero

termite.plot<-tapply(Frequency,list(loc,Taxon),sum)# Create a table with the species frequency by location
termite.plot[is.na(termite.plot)]<-0 # Species non detected in a given location are real zeros
termite.plot<-termite.plot[,colSums(termite.plot)>0] # Trim species with zero occurrences

#check if the environmental and termite tables are in the same order
#cbind(rownames(termite.plot),env[,2:5]) 

termite.plot.PA<-(termite.plot>0)+0 # Abund to PA: Easier and faster than ifelse(termite.plot>0,1,0)

# Extract species feeding and defense group as vectors
TG.RFAD<-FG2[match(colnames(termite.plot),Taxon)]
DEF.RFAD<-DEF[match(colnames(termite.plot),Taxon)]

detach(iso.RFAD)

```

# Calculate pairwise trait distance between species

We created a pairwise similarity matrix between all pairs of species. The matrix was filled with 1s and 0s, representing species pairs sharing or not a particular trait. This was done independently for feeding and defense strategy. See footnote\footnote{Multiply the vector by itself creating a pairwise matrix with one species multiplied by the other.
Then subtract from the new pairwise matrix, the values of the species multiplied by itself.
If the species trait value multiplied by its own value (square) equals the species value multiplied by the value of its pair, the pair has the same trait.
Pair has the same trait, 1, otherwise 0} for explanation on the calculation.

```{r tests_for_phylogenetic_signal_in_traits}

TGn<-as.integer(TG.RFAD)# Transform the names of feeding groups in numbers
names(TGn)<-colnames(termite.plot.PA)# Assign the names of species to the vector of feeding groups

TG.dist<-as.dist((((TGn)%*%t(TGn))-(TGn)^2)!=0)#See footnote

DEFn<-as.integer(DEF.RFAD)# Transform the names of feeding groups in numbers
names(DEFn)<-colnames(termite.plot.PA)# Assign the names of species to the vector of feeding groups

DEF.dist<-as.dist((((DEFn)%*%t(DEFn))-(DEFn)^2)!=0)#See footnote
```

# Calculate phylogenetic distance and rearrange columns to match species names in the data

Use the cophenetic function from the ape package to calculate pairwise phylogenetic distances between species. Then match the species in this matrix with those present in the species x matrix. Finally, transform the pairwise matrix in a dist object (triangular matrix).


```{r tests_for_phylogenetic_signal_in_traits_cont}

phylo.dist<-cophenetic(termite.phylo)
phylo.dist<-as.dist(phylo.dist[match(colnames(termite.plot.PA),colnames(phylo.dist)),
	match(colnames(termite.plot.PA),colnames(phylo.dist))])

```

# Calculates Moran's *I* and mantel correlograms

```{r moransI_manteltests,cache=TRUE}

Moran.I(TGn,as.matrix(phylo.dist))
Moran.I(DEFn,as.matrix(phylo.dist))

TG.mantel<-mantel.correlog(TG.dist,phylo.dist)
DEF.mantel<-mantel.correlog(DEF.dist,phylo.dist)

TG.mantel
DEF.mantel
```
\newpage

# Plot trait distances against phylogenetic distance

```{r plot_traitdistxphylodis, fig.height=6,fig.width=12,fig.cap="trait distanc against phylodis"}

par(mfrow=c(1,2),mar=c(5,2,4,2)+.1,oma=c(0,3,0,0),xpd=NA) #set graphical parameter space

plot(1-tapply(as.vector(TG.dist),as.vector(phylo.dist),mean)~sort(unique(phylo.dist)),
ylab="Probability of shared trait",xlab="Phylogenetic distance",cex.lab=1.5,pch=21,bg=1)

c1<-coef(glm(y~x,family=binomial(link="log"),data=data.frame(y=1-tapply(as.vector(TG.dist),
as.vector(phylo.dist),mean),x=sort(unique(phylo.dist)))))#line fit using glm

points(exp(cbind(1,seq(2,24,.1))%*%c1)~x,data=data.frame(x=seq(2,24,.1)),type="l")#plot line

title("A)",adj=0,cex.main=1.5)


plot(1-tapply(as.vector(DEF.dist),as.vector(phylo.dist),mean)~sort(unique(phylo.dist)),
ylab="",xlab="Phylogenetic distance",cex.lab=1.5,pch=21,bg=1)

c1<-coef(glm(y~x,family=binomial(link="log"),start = c(a=0,b=-.5),data=data.frame(y=1-tapply(
as.vector(DEF.dist),as.vector(phylo.dist),mean),x=sort(unique(phylo.dist)))))#line fit using glm

points(exp(cbind(1,seq(2,24,.1))%*%c1)~x,data=data.frame(x=seq(2,24,.1)),type="l")#plot line

title("B)",adj=0,cex.main=1.5)
```

\newpage

# Plot mantel correlograms

```{r mantel_correlograms, fig.height=6,fig.width=12}

par(mfrow=c(1,2),oma=c(0,1,0,0))#set graphical parameter space

plot(TG.mantel$mantel.res[,3]~TG.mantel$mantel.res[,1],pch=21,bg=TG.mantel$mantel.res[,5]<0.05,
xlim=c(2,12),type="o",ylab="Mantel correlation",xlab="Phylogenetic distance",cex.lab=1.5)
abline(h=0,lty=2,col=2)

title("A)",adj=0,cex.main=2)

plot(DEF.mantel$mantel.res[,3]~DEF.mantel$mantel.res[,1],pch=21,bg=DEF.mantel$mantel.res[,5]<0.05,
xlim=c(2,12),type="o",ylab="Mantel correlation",xlab="Phylogenetic distance",cex.lab=1.5)
abline(h=0,lty=2,col=2)

title("B)",adj=0,cex.main=2)


```

\newpage

# Create response variables

In the following lines, there are the code to create all the response variables used in our study. In the following sections, individual regressions of these variables against the predictor variables present in the environmental dataset are performed.

```{r create_response_variables, cache=TRUE,message=FALSE,results='hide'}

# Overall diversity

termite.N<-rowSums(termite.plot)
termite.S<-rowSums(termite.plot>0)

termite.N.w<-rowSums(termite.plot[,TG.RFAD=="W"])
termite.S.w<-rowSums(termite.plot[,TG.RFAD=="W"]>0)

termite.N.s<-rowSums(termite.plot[,TG.RFAD=="S"])
termite.S.s<-rowSums(termite.plot[,TG.RFAD=="S"]>0)

# Sorensen simiarity matrices (triangular matrices)

sor.dist<-vegdist(termite.plot,"bray",binary = T)
sor.dist.w<-vegdist(termite.plot[,TG.RFAD=="W"],"bray",binary = T)
sor.dist.s<-vegdist(termite.plot[,TG.RFAD=="S"],"bray",binary = T)

# Total phylodiversity (PD, MPD, and MNTD)

ses.pd<-ses.pd(termite.plot[,],termite.phylo,null.model = "taxa.labels")
ses.mpd<-ses.mpd(termite.plot[,],cophenetic(termite.phylo),null.model = "taxa.labels",
abundance.weighted=TRUE)
ses.mntd<-ses.mntd(termite.plot[,],cophenetic(termite.phylo),null.model = "taxa.labels",
abundance.weighted=TRUE)

# Wood feeders phylodiversity (PD, MPD, and MNTD)

ses.pd.w<-ses.pd(termite.plot[,TG.RFAD=="W"],termite.phylo,null.model = "taxa.labels")
ses.mpd.w<-ses.mpd(termite.plot[,TG.RFAD=="W"],cophenetic(termite.phylo),null.model = "taxa.labels",
abundance.weighted=TRUE)
ses.mntd.w<-ses.mntd(termite.plot[,TG.RFAD=="W"],cophenetic(termite.phylo),null.model = "taxa.labels",
abundance.weighted=TRUE)

# Soil feeders phylodiversity (PD, MPD, and MNTD)

ses.pd.s<-ses.pd(termite.plot[,TG.RFAD=="S"],termite.phylo,null.model = "taxa.labels")
ses.mpd.s<-ses.mpd(termite.plot[,TG.RFAD=="S"],cophenetic(termite.phylo),null.model = "taxa.labels",
abundance.weighted=TRUE)
ses.mntd.s<-ses.mntd(termite.plot[,TG.RFAD=="S"],cophenetic(termite.phylo),null.model = "taxa.labels",
abundance.weighted=TRUE)

# Total phylocomposition (Unifrac,PhyloSorensen, and betaMNTD [not used])

ufrac.dist<-unifrac(termite.plot,termite.phylo)
psor.dist<-phylosor(termite.plot,termite.phylo)
betaMNTD.dist<-comdistnt(termite.plot,cophenetic(termite.phylo))

# Wood-feeders phylocomposition (Unifrac,PhyloSorensen, and betaMNTD [not used])

ufrac.dist.w<-unifrac(termite.plot[,TG.RFAD=="W"],termite.phylo)
psor.dist.w<-phylosor(termite.plot[,TG.RFAD=="W"],termite.phylo)
betaMNTD.dist.w<-comdistnt(termite.plot[,TG.RFAD=="W"],cophenetic(termite.phylo))

# Soil-feeders phylocomposition (Unifrac,PhyloSorensen, and betaMNTD [not used])

ufrac.dist.s<-unifrac(termite.plot[,TG.RFAD=="S"],termite.phylo)
psor.dist.s<-phylosor(termite.plot[,TG.RFAD=="S"],termite.phylo)
betaMNTD.dist.s<-comdistnt(termite.plot[,TG.RFAD=="S"],cophenetic(termite.phylo))


# Functional diversity

names(TG.RFAD)<-colnames(termite.plot)
names(DEF.RFAD)<-colnames(termite.plot)

TG.fdis<-fdisp(TG.dist,termite.plot)$FDis
DEF.fdis<-fdisp(DEF.dist,termite.plot)$FDis


#standardied effect size for trait diversity

nrep=999

TG.fdis.sim<-matrix(nrow=nrow(termite.plot),ncol=nrep)
TG.dist.null<-TG.dist

DEF.fdis.sim<-matrix(nrow=nrow(termite.plot),ncol=nrep)
DEF.dist.null<-DEF.dist

for(i in 1:nrep){

TG.dist.null[]<-sample(TG.dist)
DEF.dist.null[]<-sample(DEF.dist)

TG.fdis.sim[,i]<-fdisp(TG.dist.null,termite.plot)$FDis
DEF.fdis.sim[,i]<-fdisp(DEF.dist.null,termite.plot)$FDis
	
	}

TG.fdis.ses<-(TG.fdis-rowMeans(TG.fdis.sim))/apply(TG.fdis.sim,1,sd)
DEF.fdis.ses<-(DEF.fdis-rowMeans(DEF.fdis.sim))/apply(DEF.fdis.sim,1,sd)


# Functional composition

#anova.cca(cca(t(aggregate(t(termite.plot),by = list(TG.RFAD),sum)[,-1])~.,data=env.RFAD[,c("ant.pred.N","P")]),by = "terms")	
#anova.cca(cca(t(aggregate(t(termite.plot),by = list(DEF.RFAD),sum)[,-1])~.,data=env.RFAD[,c("ant.pred.N","P")]),by = "terms")	

TG.pcoa<-prcomp(t(aggregate(t(termite.plot),by = list(TG.RFAD),sum)[,-1]))$x[,1]
DEF.pcoa<-prcomp(t(aggregate(t(termite.plot),by = list(DEF.RFAD),sum)[,-1]))$x[,1]

TG.nmds<-scores(metaMDS(t(aggregate(t(termite.plot),by = list(TG.RFAD),sum)[,-1])))[,1]
DEF.nmds<-scores(metaMDS(t(aggregate(t(termite.plot),by = list(DEF.RFAD),sum)[,-1])))[,1]


# Separate nestedness and turnover on functional composition (Baselga 2010)
# Not performed because all groups are represented in all sites (dissimilarity using sor = 0)

#TG.sim.dist<-beta.sim(t(aggregate(t(termite.plot),by = list(TG.RFAD),sum)[,-1]))
#TG.nes.dist<-beta.nes(t(aggregate(t(termite.plot),by = list(TG.RFAD),sum)[,-1]))

#nes.dist.w<-beta.nes(termite.plot[,TG.RFAD=="W"])
#nes.dist.s<-beta.nes(termite.plot[,TG.RFAD=="S"])

```


# Summarize composition and phylocomposition in PCoA axes

Similar to Duarte et al. (2011), we summarized the phylogenetic dissimilarity matrices in PCoA axes, representing phylogenetic composition. This was performed to all indexes used in our study (Sorensen, PhyloSor, and UniFrac).


```{r pcoas_comp_and_phylocomp,results='hide'}

pcoa.sor<-scores(cmdscale(sor.dist))

pcoa.psor<-scores(cmdscale(psor.dist))
pcoa.ufrac<-scores(cmdscale(ufrac.dist))

pcoa.sor.w<-scores(cmdscale(sor.dist.w))

pcoa.psor.w<-scores(cmdscale(psor.dist.w))
pcoa.ufrac.w<-scores(cmdscale(ufrac.dist.w))

pcoa.sor.s<-scores(cmdscale(sor.dist.s))
pcoa.psor.s<-scores(cmdscale(psor.dist.s))
pcoa.ufrac.s<-scores(cmdscale(ufrac.dist.s))
```


```{r nmds_comp_and_phylocomp,echo=FALSE,eval=FALSE}


nmds.sor<-scores(metaMDS(sor.dist))

nmds.psor<-scores(metaMDS(psor.dist))
nmds.ufrac<-scores(metaMDS(ufrac.dist))

nmds.sor.w<-scores(metaMDS(sor.dist.w))

nmds.psor.w<-scores(metaMDS(psor.dist.w))
nmds.ufrac.w<-scores(metaMDS(ufrac.dist.w))

nmds.sor.s<-scores(metaMDS(sor.dist.s))
nmds.psor.s<-scores(metaMDS(psor.dist.s))
nmds.ufrac.s<-scores(metaMDS(ufrac.dist.s))

```


\newpage

# Regression analyses by response variable

For each response variable, we run a multiple regression against the predictor variables. P-values using t and F statistics were similar, but we based our results on the F-statistics results. This answer the question: *Does the inclusion of a given variable significantly increases the explanatory power of the model?*
You can see a summary of these results as a table in the end of the document

```{r Results.Ass_diversity_env_var_simple,echo=FALSE,eval=FALSE}

##Overall

summary(lm(ses.pd$pd.obs.z~env.RFAD$ant.pred.N))
summary(lm(ses.pd$pd.obs.z~env.RFAD$ant.nonpred.N))
summary(lm(ses.pd$pd.obs.z~env.RFAD$P))
summary(lm(ses.pd$pd.obs.z~env.RFAD$N.percentage))
summary(lm(ses.pd$pd.obs.z~env.RFAD$Clay))

summary(lm(ses.mpd$mpd.obs.z~env.RFAD$ant.pred.N))
summary(lm(ses.mpd$mpd.obs.z~env.RFAD$P))
summary(lm(ses.mpd$mpd.obs.z~env.RFAD$N.percentage))
summary(lm(ses.mpd$mpd.obs.z~env.RFAD$Clay))

summary(lm(ses.mntd$mntd.obs.z~env.RFAD$ant.pred.N))
summary(lm(ses.mntd$mntd.obs.z~env.RFAD$P))
summary(lm(ses.mntd$mntd.obs.z~env.RFAD$N.percentage))
summary(lm(ses.mntd$mntd.obs.z~env.RFAD$Clay))

# Wood feeders

summary(lm(ses.pd.w$ntaxa~env.RFAD$ant.pred.N))
summary(lm(ses.pd.w$ntaxa~env.RFAD$P))
summary(lm(ses.pd.w$ntaxa~env.RFAD$N.percentage))
summary(lm(ses.pd.w$ntaxa~env.RFAD$Clay))


summary(lm(ses.pd.w$pd.obs.z~env.RFAD$ant.pred.N))
summary(lm(ses.pd.w$pd.obs.z~env.RFAD$P))
summary(lm(ses.pd.w$pd.obs.z~env.RFAD$N.percentage))
summary(lm(ses.pd.w$pd.obs.z~env.RFAD$Clay))

summary(lm(ses.mpd.w$mpd.obs.z~env.RFAD$ant.pred.N))
summary(lm(ses.mpd.w$mpd.obs.z~env.RFAD$P))
summary(lm(ses.mpd.w$mpd.obs.z~env.RFAD$N.percentage))
summary(lm(ses.mpd.w$mpd.obs.z~env.RFAD$Clay))

summary(lm(ses.mntd.w$mntd.obs.z~env.RFAD$ant.pred.N))
summary(lm(ses.mntd.w$mntd.obs.z~env.RFAD$P))
summary(lm(ses.mntd.w$mntd.obs.z~env.RFAD$N.percentage))
summary(lm(ses.mntd.w$mntd.obs.z~env.RFAD$Clay))

# Soil feeders

summary(lm(ses.pd.s$pd.obs.z~env.RFAD$ant.pred.N))
summary(lm(ses.pd.s$pd.obs.z~env.RFAD$P))
summary(lm(ses.pd.s$pd.obs.z~env.RFAD$N.percentage))
summary(lm(ses.pd.s$pd.obs.z~env.RFAD$Clay))

summary(lm(ses.mpd.s$mpd.obs.z~env.RFAD$ant.pred.N))
summary(lm(ses.mpd.s$mpd.obs.z~env.RFAD$P))
summary(lm(ses.mpd.s$mpd.obs.z~env.RFAD$N.percentage))
summary(lm(ses.mpd.s$mpd.obs.z~env.RFAD$Clay))

summary(lm(ses.mntd.s$mntd.obs.z~env.RFAD$ant.pred.N))
summary(lm(ses.mntd.s$mntd.obs.z~env.RFAD$P))
summary(lm(ses.mntd.s$mntd.obs.z~env.RFAD$N.percentage))
summary(lm(ses.mntd.s$mntd.obs.z~env.RFAD$Clay))

```

## Overall PD

```{r PD_multiple}
summary(M1<-lm(ses.pd$pd.obs.z~ant.pred.N+P+N.percentage+Vegetation.structure,data=env.RFAD))
anova(M1)


```
## Overall MPD
```{r MPD_multiple}
summary(M1<-lm(ses.mpd$mpd.obs.z~ant.pred.N+P+N.percentage+Vegetation.structure,data=env.RFAD))
anova(M1)
```
## Overall MNTD
```{r NMTD_multiple}
summary(M1<-lm(ses.mntd$mntd.obs.z~ant.pred.N+P+N.percentage+Vegetation.structure,data=env.RFAD))
anova(M1)
```

## Overall Sorensen
```{r sor_multiple}
summary(M1<-lm(pcoa.sor[,1]~ant.pred.N+P+N.percentage+Vegetation.structure,data=env.RFAD))
anova(M1)
```

## Overall PhyloSorensen
```{r psor_multiple}
summary(M1<-lm(pcoa.psor[,1]~ant.pred.N+P+N.percentage+Vegetation.structure,data=env.RFAD))
anova(M1)
```

## Overall Unifrac
```{r ufrac_multiple}
summary(M1<-lm(pcoa.ufrac[,1]~ant.pred.N+P+N.percentage+Vegetation.structure,data=env.RFAD))
anova(M1)
```

## Wood feeders PD
```{r PDw_multiple}
summary(M1<-lm(ses.pd.w$pd.obs.z~ant.pred.N+P+N.percentage+Vegetation.structure,data=env.RFAD))
anova(M1)
```
## Wood feeders MPD
```{r MPDw_multiple}
summary(M1<-lm(ses.mpd.w$mpd.obs.z~ant.pred.N+P+N.percentage+Vegetation.structure,data=env.RFAD))
anova(M1)
```
## Wood feeders MNTD
```{r MNTDw_multiple}
summary(M1<-lm(ses.mntd.w$mntd.obs.z~ant.pred.N+P+N.percentage+Vegetation.structure,data=env.RFAD))
anova(M1)
```

## Wood feeders Sorensen
```{r sorw_multiple}
summary(M1<-lm(pcoa.sor.w[,1]~ant.pred.N+P+N.percentage+Vegetation.structure,data=env.RFAD))
anova(M1)
```

## Wood feeders PhyloSorensen
```{r psorw_multiple}
summary(M1<-lm(pcoa.psor.w[,1]~ant.pred.N+P+N.percentage+Vegetation.structure,data=env.RFAD))
anova(M1)
```

## Wood feeders Unifrac
```{r ufracw_multiple}
summary(M1<-lm(pcoa.ufrac.w[,1]~ant.pred.N+P+N.percentage+Vegetation.structure,data=env.RFAD))
anova(M1)
```


## Soil feeders PD
```{r PDs_multiple}
summary(M1<-lm(ses.pd.s$pd.obs.z~ant.pred.N+P+N.percentage+Vegetation.structure,data=env.RFAD))
anova(M1)
```
## Soil feeders MPD

```{r MPDs_multiple}
summary(M1<-lm(ses.mpd.s$mpd.obs.z~ant.pred.N+P+N.percentage+Vegetation.structure,data=env.RFAD))
anova(M1)
```

## Soil feeders MNTD

```{r MNTDs_multiple}
summary(M1<-lm(ses.mntd.s$mntd.obs.z~ant.pred.N+P+N.percentage+Vegetation.structure,data=env.RFAD))
anova(M1)

```
## Soil feeders Sorensen
```{r sors_multiple}
summary(M1<-lm(pcoa.sor.s[,1]~ant.pred.N+P+N.percentage+Vegetation.structure,data=env.RFAD))
anova(M1)
```

## Soil feeders PhyloSorensen
```{r psors_multiple}
summary(M1<-lm(pcoa.psor.s[,1]~ant.pred.N+P+N.percentage+Vegetation.structure,data=env.RFAD))
anova(M1)
```

## Soil feeders Unifrac
```{r ufracs_multiple}
summary(M1<-lm(pcoa.ufrac.s[,1]~ant.pred.N+P+N.percentage+Vegetation.structure,data=env.RFAD))
anova(M1)
```

\newpage

# Plot phylogenetic diversity against *P*

One of the strongest findings in this study was the strong relation of phylogenetic diversity metrics with *P* content.

```{r plots, fig.height=4,fig.width=12,results='hold'}

par(mfrow=c(1,3),mar=c(5,2,4,2)+.1,oma=c(0,3,0,0),xpd=NA)

plot(ses.mntd$mntd.obs.z~env.RFAD$P,cex.lab=1.5,xlab="P content", ylab= "Phylogenetic diversity",pch=21,col=0,bg=1,cex=1.2,ylim=c(-2.5,2.5))

#c1<-coef(nls(y~a+b*exp(c*x),start=list(a=-1,b=1,c=-0.2),data=data.frame(y=ses.mntd$mntd.obs.z,x=env.RFAD$P)))
#points((c1[1]+c1[2]*exp(c1[3]*x))~x,data=data.frame(x=seq(-1,2.5,.1)),type="l")

title("A)",adj=0,cex.main=1.5)

plot(ses.mntd.w$mntd.obs.z~env.RFAD$P,cex.lab=1.5,xlab="P content", ylab= "",pch=21,col=0,bg=1,cex=1.2,ylim=c(-2.5,2.5))

#c1<-coef(nls(y~a+b*exp(c*x),start=list(a=-3,b=5,c=-0.8),data=data.frame(y=ses.mntd.w$mntd.obs.z,x=env.RFAD$P)))
#points((c1[1]+c1[2]*exp(c1[3]*x))~x,data=data.frame(x=seq(-1,2.5,.1)),type="l")

title("B)",adj=0,cex.main=1.5)

plot(ses.mntd.s$mntd.obs.z~env.RFAD$P,cex.lab=1.5,xlab="P content", ylab= "",pch=21,col=0,bg=1,cex=1.2,ylim=c(-2.5,2.5))

#c1<-coef(nls(y~a+b*exp(c*x),start=list(a=1,b=1,c=0.2),data=data.frame(y=ses.mntd.s$mntd.obs.z,x=env.RFAD$P)))
#points((c1[1]+c1[2]*exp(c1[3]*x))~x,data=data.frame(x=seq(-1,2.5,.1)),type="l")


title("C)",adj=0,cex.main=1.5)

```

```{r, echo=FALSE,eval=FALSE}

# Plot termite phylogeny showing the distribution in trophic groups

op<-par(no.readonly=TRUE)

plot(termite.phylo, show.tip.label=F,plot=TRUE)

colors<-as.numeric(TG.RFAD)[match(termite.phylo$tip.label,colnames(termite.plot))]
colors[is.na(colors)]<-4

tiplabels(pch=22, col=0, bg=colors, cex=1.5,adj=.7)

###

#Plot the results for the total using MNTD (strongest relation)

par(op)

layout(matrix(c(1,1,2,1,1,2,1,1,3,1,1,3),3,4))

par(mar=c(5,5,4,2))
plot(ses.mntd$mntd.obs.z~env.RFAD$P,cex.lab=1.5,xlab="P content", ylab= "Phylogenetic diversity",pch=21,col=0,bg=1,cex=1.2)

title(main="A)",adj=0,cex.main=2)

par(mar=c(1,2,2,2))
plot(termite.phylo, show.tip.label=F,plot=TRUE)

#col.lowp<-termite.plot.PA[order(env.RFAD$P)==1,match(termite.phylo$tip.label,colnames(termite.plot))]

col.highmntd<-termite.plot.PA[order(ses.mntd$mntd.obs.z,decreasing = TRUE)==1,match(termite.phylo$tip.label,colnames(termite.plot))]

tiplabels(pch=22, col=col.highmntd, bg=colors*col.highmntd, cex=1,adj=1)

title(main="B)",adj=0,cex.main=2)

plot(termite.phylo, show.tip.label=F,plot=TRUE)

#col.highp<-termite.plot.PA[order(env.RFAD$P,decreasing = TRUE)==1,match(termite.phylo$tip.label,colnames(termite.plot))]

col.lowmntd<-termite.plot.PA[order(ses.mntd$mntd.obs.z)==1,match(termite.phylo$tip.label,colnames(termite.plot))]

tiplabels(pch=22, col=col.lowmntd, bg=colors*col.lowmntd, cex=1,adj=1)

title(main="C)",adj=0,cex.main=2)

par(op)

```

\newpage

# Diagrams: spp. presence by phylo and predictor variable

In these diagrams, species were color coded by their feeding and defense strategies. The species were ordered by their position in the phylogeny and the locations where they were found were ordered by *P* and *ant predator density*


```{r poncho1, fig.height=6,fig.width=5}

palette(c("#004CFFFF","black", "#00FF4DFF","#00E5FFFF","tomato","yellow","grey"))

poncho(termite.plot.PA,phy=termite.phylo,env=env.RFAD$P,col=as.integer(TG.RFAD)+1,
	file="phyloxPA",ylab.bottom = expression("P "(mg/dm^3)),xlab="Sampling units",
	col.gradient = "tomato",border="black")

```

\newpage

```{r poncho2, fig.height=6,fig.width=5}

poncho(termite.plot.PA,phy=termite.phylo,env=env.RFAD$ant.pred.N,col=as.integer(DEF.RFAD),
	ylab.bottom = "Ant density",xlab="Sampling units",col.gradient = "tomato",border="black")

palette("default")

```


```{r, eval=FALSE,echo=FALSE}

par(mar=c(0,0,0,0),oma=c(0,0,0,0))

layout(matrix(c(1,1,1,2,1,1,1,2),4,2))

sp.data<-termite.plot.PA
phy<-termite.phylo
env<-10*(2+env.RFAD$P)
xlab=expression("P content" (mg/dm^3))

sp.color<-as.integer(TG.RFAD)
z<-(t(t(sp.data)*sp.color))[order(env),match(phy$tip.label,colnames(sp.data))]

plot(phy, show.tip.label=F,plot=TRUE,no.margin = TRUE,x.lim=25)

space.x=(25-13.5)/31
space.y=1

x1=((z>0)*seq(13.5,25-space.x,length=nrow(z)))[(!is.na(z))&z>0]
y1=(t(t(z>0)*((1:length(phy$tip.label))-0.5)))[(!is.na(z))&z>0]

segments(13.4,1:length(phy$tip.label),25.25,1:length(phy$tip.label),lty=3,col="dark grey")

rect(x1,y1,x1+space.x,y1+space.y,col=c(1,4,3)[z[(!is.na(z))&z>0]])

plot(phy,plot=FALSE,no.margin = TRUE,x.lim=25)

rect(seq(13.5,25-space.x,length=nrow(z)),35,seq(13.5,25-space.x,length=nrow(z))+space.x,sort(env+35),col="dark grey")

text(25-(25-13.5)/2,15,xlab,cex=2)

segments(c(13.2,13.2,13.2),c(35,35,ceiling(max(env)+35)),c(13.2,13.2-.2,13.2-.2),c(ceiling(max(env)+35),35,ceiling(max(env)+35)))
text(13.2-.3,c(35,ceiling(max(env)+35)),c(0,ceiling(max(env))),adj=1)


```




```{r regr_first_phylocomp_axis_x_en,echo=FALSE,eval=FALSE}

summary(lm(nmds.sor[,1]~env.RFAD$ant.pred.N))
summary(lm(nmds.sor[,1]~env.RFAD$P))
summary(lm(nmds.sor[,1]~env.RFAD$N.percentage))
summary(lm(nmds.sor[,1]~env.RFAD$Clay))

summary(lm(nmds.sor.w[,1]~env.RFAD$ant.pred.N))
summary(lm(nmds.sor.w[,1]~env.RFAD$P))
summary(lm(nmds.sor.w[,1]~env.RFAD$N.percentage))
summary(lm(nmds.sor.w[,1]~env.RFAD$Clay))

summary(lm(nmds.sor.s[,1]~env.RFAD$ant.pred.N))
summary(lm(nmds.sor.s[,1]~env.RFAD$P))
summary(lm(nmds.sor.s[,1]~env.RFAD$N.percentage))
summary(lm(nmds.sor.s[,1]~env.RFAD$Clay))

###

summary(lm(nmds.psor[,1]~env.RFAD$ant.pred.N))
summary(lm(nmds.psor[,1]~env.RFAD$P))
summary(lm(nmds.psor[,1]~env.RFAD$N.percentage))
summary(lm(nmds.psor[,1]~env.RFAD$Clay))

summary(lm(nmds.psor.w[,1]~env.RFAD$ant.pred.N))
summary(lm(nmds.psor.w[,1]~env.RFAD$P))
summary(lm(nmds.psor.w[,1]~env.RFAD$N.percentage))
summary(lm(nmds.psor.w[,1]~env.RFAD$Clay))

summary(lm(nmds.psor.s[,1]~env.RFAD$ant.pred.N))
summary(lm(nmds.psor.s[,1]~env.RFAD$P))
summary(lm(nmds.psor.s[,1]~env.RFAD$N.percentage))
summary(lm(nmds.psor.s[,1]~env.RFAD$Clay))

###

summary(lm(nmds.ufrac[,1]~env.RFAD$ant.pred.N))
summary(lm(nmds.ufrac[,1]~env.RFAD$P))
summary(lm(nmds.ufrac[,1]~env.RFAD$N.percentage))
summary(lm(nmds.ufrac[,1]~env.RFAD$Clay))

summary(lm(nmds.ufrac.w[,1]~env.RFAD$ant.pred.N))
summary(lm(nmds.ufrac.w[,1]~env.RFAD$P))
summary(lm(nmds.ufrac.w[,1]~env.RFAD$N.percentage))
summary(lm(nmds.ufrac.w[,1]~env.RFAD$Clay))

summary(lm(nmds.ufrac.s[,1]~env.RFAD$ant.pred.N))
summary(lm(nmds.ufrac.s[,1]~env.RFAD$P))
summary(lm(nmds.ufrac.s[,1]~env.RFAD$N.percentage))
summary(lm(nmds.ufrac.s[,1]~env.RFAD$Clay))

```

```{r funct_diver,echo=FALSE,results='hide'}

summary(lm(TG.fdis.ses~env.RFAD$ant.pred.N))
summary(lm(TG.fdis.ses~env.RFAD$P))
summary(lm(TG.fdis.ses~env.RFAD$N.percentage))
summary(lm(TG.fdis.ses~env.RFAD$Vegetation.structure))

summary(lm(DEF.fdis.ses~env.RFAD$ant.pred.N))
summary(lm(DEF.fdis.ses~env.RFAD$P))
summary(lm(DEF.fdis.ses~env.RFAD$N.percentage))
summary(lm(DEF.fdis.ses~env.RFAD$Vegetation.structure))

```

```{r funct_comp,echo=FALSE,results='hide'}

summary(lm(TG.nmds~env.RFAD$ant.pred.N))
summary(lm(TG.nmds~env.RFAD$P))
summary(lm(TG.nmds~env.RFAD$N.percentage))
summary(lm(TG.nmds~env.RFAD$Vegetation.structure))

summary(lm(DEF.nmds~env.RFAD$ant.pred.N))
summary(lm(DEF.nmds~env.RFAD$P))
summary(lm(DEF.nmds~env.RFAD$N.percentage))
summary(lm(DEF.nmds~env.RFAD$Vegetation.structure))

##

summary(lm(TG.pcoa~env.RFAD$ant.pred.N))
summary(lm(TG.pcoa~env.RFAD$P))
summary(lm(TG.pcoa~env.RFAD$N.percentage))
summary(lm(TG.pcoa~env.RFAD$Vegetation.structure))

summary(lm(DEF.pcoa~env.RFAD$ant.pred.N))
summary(lm(DEF.pcoa~env.RFAD$P))
summary(lm(DEF.pcoa~env.RFAD$N.percentage))
summary(lm(DEF.pcoa~env.RFAD$Vegetation.structure))

```


```{r table1, echo=FALSE,results='hide'}

#####

table1<-lapply(list(
	
	ses.pd$pd.obs.z,
	ses.mpd$mpd.obs.z,
	ses.mntd$mntd.obs.z,
	pcoa.sor[,1],
	pcoa.psor[,1],
	pcoa.ufrac[,1],

	ses.pd.w$pd.obs.z,
	ses.mpd.w$mpd.obs.z,
	ses.mntd.w$mntd.obs.z,
	pcoa.sor.w[,1],
	pcoa.psor.w[,1],
	pcoa.ufrac.w[,1],
	
	ses.pd.s$pd.obs.z,
	ses.mpd.s$mpd.obs.z,
	ses.mntd.s$mntd.obs.z,
	pcoa.sor.s[,1],
	pcoa.psor.s[,1],
	pcoa.ufrac.s[,1],
	
	TG.fdis,
	DEF.fdis,
	
	TG.pcoa,
	DEF.pcoa
		
	),function(x){
	
		
	c(m.var=coef(M1<-lm(x~ant.pred.N+P+N.percentage+Vegetation.structure,data=env.RFAD))[-1],
		m.df=summary(M1)$df[2],
		m.F=summary(M1)$fstatistic[1],
		m.var.rsq=summary(M1)$adj.r.squared*100,
		m.pval.p=anova(M1)[-5,"Pr(>F)"],
		m.pval.t=pf(summary(M1)$fstatistic[1], summary(M1)$fstatistic[2], summary(M1)$fstatistic[3],lower.tail = FALSE),		
		var=c(coef(update(M1,~ant.pred.N))[2],coef(update(M1,~P))[2],coef(update(M1,~N.percentage))[2],coef(update(M1,~Clay))[2]),
		pval=drop1(M1,test="F")[-1,"Pr(>F)"], 
		var.rsq=100*c(summary(update(M1,~ant.pred.N))$r.squared,summary(update(M1,~P))$r.squared,summary(update(M1,~N.percentage))$r.squared,summary(update(M1,~Clay))$r.squared))
	})

table1<-do.call(rbind,table1)

rownames(table1)<-c(rep(c("PD","MPD","MNTD","PCoA.sor","PCoA.psor","PCoA.ufrac"),3),"TG.FD","DEF.FD","PCoA.TG","PCoA.DEF")# for all, wood, soil

write.csv((round(table1,3)),"Table1.raw.csv")	

table1[,7]<-ifelse(table1[,7]<0,0,table1[,7])

pval.sym<-c("***"=0,"**"=0.001,"*"=0.01,"+"=0.05,0.1,1)

pval<-names(pval.sym)[as.integer(
	cut(table1[,grep("pval",colnames(table1))],breaks = pval.sym,include.lowest = T)
	)]

table1.f<-round(table1,3)

table1.f[,grep("var",colnames(table1.f))]<-paste(table1.f[,grep("var",colnames(table1.f))],pval,sep="")
table1.f<-table1.f[,-grep("pval",colnames(table1.f))]

colnames(table1.f)<-gsub("var.","",colnames(table1.f))

write.csv(table1.f,"Table1.final.csv")

table1.1<-table1.f[,1:7]

colnames(table1.1)<-c("N ant pred","P","N","N trees","df","F","r2")

table1.1<-ifelse(table1.1=="0","0.000",table1.1)

write.csv(table1.1,"Table1.1.final.csv")

```

\newpage


# Summary of results in Table 1
 
* Code for generating table not shown. You can check the results in the Regressions section 

```{r show_table1,echo=FALSE}

data.frame(response=rownames(table1.1),table1.1)

```



```{r plot_phylocomp_x_env,echo=FALSE,eval=FALSE}

plot(pcoa.psor[,1]~env.RFAD$ant.pred.N)
plot(pcoa.psor[,1]~env.RFAD$P)
plot(pcoa.psor[,1]~env.RFAD$N.percentage)
plot(pcoa.psor[,1]~env.RFAD$Clay)

plot(pcoa.psor.w[,1]~env.RFAD$ant.pred.N)
plot(pcoa.psor.w[,1]~env.RFAD$P)
plot(pcoa.psor.w[,1]~env.RFAD$N.percentage)
plot(pcoa.psor.w[,1]~env.RFAD$Clay)

plot(pcoa.psor.s[,1]~env.RFAD$ant.pred.N)
plot(pcoa.psor.s[,1]~env.RFAD$P)
plot(pcoa.psor.s[,1]~env.RFAD$N.percentage)
plot(pcoa.psor.s[,1]~env.RFAD$Clay)

plot(pcoa.ufrac[,1]~env.RFAD$ant.pred.N)
plot(pcoa.ufrac[,1]~env.RFAD$P)
plot(pcoa.ufrac[,1]~env.RFAD$N.percentage)
plot(pcoa.ufrac[,1]~env.RFAD$Clay)

plot(pcoa.ufrac.w[,1]~env.RFAD$ant.pred.N)
plot(pcoa.ufrac.w[,1]~env.RFAD$P)
plot(pcoa.ufrac.w[,1]~env.RFAD$N.percentage)
plot(pcoa.ufrac.w[,1]~env.RFAD$Clay)

plot(pcoa.ufrac.s[,1]~env.RFAD$ant.pred.N)
plot(pcoa.ufrac.s[,1]~env.RFAD$P)
plot(pcoa.ufrac.s[,1]~env.RFAD$N.percentage)
plot(pcoa.ufrac.s[,1]~env.RFAD$Clay)

##

```




